### 2020-05-12 writeup/proof_transfer_ex.tex
sec 2 edits
proof of prop 3.1

From \cite{WZR20}, we know that we need to explicitly restrict the capacity $r$ of $B$ so that there is transfer between the two tasks.
for the rest of the section, we shall consider the case when $r=1$ we are considering the case of two tasks.
Here, equation \eqref{eq_mtl} simplifies to the following
\[ f(B; w_1, w_2) = \bignorm{X_1 B w_1 - Y_1}^2 + \bignorm{X_2 B w_2 - Y_2}^2, \]
where $B\in\real^p$ and $w_1, w_2$ are both real numbers.
To solve the above, suppose that $w_1, w_2$ are fixed, by local optimality, we solve $B$ as
\begin{align*}
	\hat{B}(w) &= (w_1^2 X_1^{\top}X_1 + w_2^2 X_2^{\top}X_2)^{-1} (w_1 X_1^{\top}Y_1 + w_2 X_2^{\top}Y_2) \\
	&= \frac{1}{w_2} (w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} (w X_1^{\top}Y_1 + X_2^{\top}Y_2) \\
	&= \frac{1}{w_2}\bigbrace{\beta_t + (w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\bigbrace{X_1^{\top}X_1(w\beta_s - w^2\beta_t) + (w X_1^{\top}\varepsilon_1 + X_2^{\top}\varepsilon_2)}},
\end{align*}
where we denote $w = w_1 / w_2$.
As a remark, when $w = 1$, we recover the linear regression estimator.

\medskip
\noindent\textit{Jointly optimizing over both tasks.}
Using a validation set that is sub-sampled from the original training dataset, we get a validation loss as follows
\begin{align}
		&\val(\hat{B}; w_1, w_2) \nonumber\\
	=& n_1 \cdot \bigbrace{\bignorm{\Sigma_1^{1/2}(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_2^{\top}X_2(\beta_s - w\beta_t)}^2 + \sigma^2 \cdot \bigtr{(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1}} \nonumber \\
	+& n_2 \cdot \bigbrace{w^2\bignorm{\Sigma_2^{1/2}(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1(\beta_s - w\beta_t)}^2 + \sigma^2 \cdot \bigtr{(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}} \label{eq_val_mtl}
\end{align}
Minimizing the above loss is akin to performing multi-task training in practice.
Let $\hat{w}$ denote the minimizer of $\val(\hat{B}; w_1, w_2)$ over $w\in\real$.
We will denote $\hat{\beta}_t^{\MTL} = w_{2}\hat{B}(\hat{w})$.

Our goal is to study under model and covariate shifts, whether multi-task learning helps learn the target task better than single-task learning.
The baseline where we solve the target task with its own data is
\begin{align*}
	te(\hat{\beta}_t^{\STL}) = \sigma^2 \cdot \bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}}, \text{ where } \hat{\beta}_t^{\STL} = (X_2^{\top}X_2)^{-1} X_2^{\top}Y_2.
\end{align*}
Clearly, whether $\hat{\beta}_t^{\MTL}$ outperforms $\hat{\beta}_t^{\STL}$ depends on the covariate matrices, the difference of the task models, and the number of per-task data points.

\begin{lemma}\label{lem_minv}[\todo{ref?}]
	In the setting when $n_2 = c_2 p$ we have that as $p$ goes to infinity,
	\[ \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{p}{n_2 - p}. \]
\end{lemma}

\begin{lemma}\label{lem_cov_shift}
	In the setting of Theorem \ref{thm_model_shift}, we have with high probability $1-{\rm o}(1)$,
\begin{align}\label{lem_cov_shift_eq}
\tr ((w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2) = \frac{1}{n_1+n_2}\cdot \tr \left( \frac{1}{a_1 M^\top M + a_2} \right) +\bigo{ n^{-1/2+\epsilon}},
\end{align}
	for any constant $\epsilon>0$.
\end{lemma}
We will give the proof of this lemma in Section \ref{sec_maintools}.

@@

>>>>

From \cite{WZR20}, we know that we need to explicitly restrict the capacity $r$ of $B$ so that there is transfer between the two tasks.
for the rest of the section, we shall consider the case when $r=1$ we are considering the case of two tasks.
Here, equation \eqref{eq_mtl} simplifies to the following
\begin{align}\label{eq_mtl_2task}
	f(B; w_1, w_2) = \bignorm{X_1 B w_1 - Y_1}^2 + \bignorm{X_2 B w_2 - Y_2}^2,
\end{align}
where $B\in\real^p$ and $w_1, w_2$ are both real numbers.
To solve the above, suppose that $w_1, w_2$ are fixed, by local optimality, we solve $B$ as
\begin{align*}
	& \hat{B}(w_1, w_2) \\
	=& (w_1^2 X_1^{\top}X_1 + w_2^2 X_2^{\top}X_2)^{-1} (w_1 X_1^{\top}Y_1 + w_2 X_2^{\top}Y_2) \\
	=& \frac{1}{w_2} ((\frac{w_1}{w_2})^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} (\frac{w_1}{w_2} X_1^{\top}Y_1 + X_2^{\top}Y_2) \\
	=& \frac{1}{w_2}\bigbrace{\beta_t + ((\frac{w_1}{w_2})^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\bigbrace{X_1^{\top}X_1(\frac{w_1}{w_2}\beta_s - w^2\beta_t) + (\frac{w_1}{w_2} X_1^{\top}\varepsilon_1 + X_2^{\top}\varepsilon_2)}}.
\end{align*}
As a remark, when $w_1 = w_2 = 1$, we recover the linear regression estimator.
The advantage of using $f(B; w_1, w_2)$ is that if $\theta_1$ is a scaling of $\theta_2$, then this case can be solved optimally using equation \eqref{eq_mtl_2task} \cite{KD12}.

\textbf{Defining the multi-task learning estimator.}
Using a validation set that is sub-sampled from the original training dataset, we get a validation loss as follows
\begin{align}
		\val(\hat{B}; w_1, w_2)
	= & n_1 \cdot \bignorm{\Sigma_1^{1/2}(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_2^{\top}X_2(\beta_s - w\beta_t)}^2 \nonumber \\
		&+ n_1 \sigma^2 \cdot \bigtr{(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1} \nonumber \\
		&+ n_2 \cdot w^2\bignorm{\Sigma_2^{1/2}(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1(\beta_s - w\beta_t)}^2 \nonumber \\
		&+ n_2 \cdot \sigma^2 \cdot \bigtr{(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} \label{eq_val_mtl}
\end{align}
Let $\hat{w_1}, \hat{w_2}$ be the global minimizer of $\val(\hat{B}; w_1, w_2)$.
We will define the multi-task learning estimator for the target task as
	\[ \hat{\beta}_t^{\MTL} = \hat{w}_{2}\hat{B}(\hat{w}_1, \hat{w}_2). \]

The intuition for deriving $\hat{\beta}_t^{\MTL}$ is akin to performing multi-task training in practice.
Let $\hat{v} = \hat{w_1} / \hat{w_2}$ for the simplicity of notation.
The test loss of using $\hat{\beta}_t^{\MTL}$ for the target task is
\begin{align}
	\te(\hat{\beta}_t^{\MTL}) =&~ \hat{v}^2 \bignorm{\Sigma_2^{1/2}(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_s - \hat{v} \beta_t)}^2 \nonumber \\
			&~ + \sigma^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma_2}. \label{eq_te_mtl_2task}
\end{align}

Our goal is to study under model and covariate shifts, whether multi-task learning helps learn the target task better than single-task learning.
The baseline where we solve the target task with its own data is
\begin{align*}
	te(\hat{\beta}_t^{\STL}) = \sigma^2 \cdot \bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}}, \text{ where } \hat{\beta}_t^{\STL} = (X_2^{\top}X_2)^{-1} X_2^{\top}Y_2.
\end{align*}
%Clearly, whether $\hat{\beta}_t^{\MTL}$ outperforms $\hat{\beta}_t^{\STL}$ depends on the covariate matrices, the difference of the task models, and the number of per-task data points.

We first state several helper lemmas that can be viewed as corollaries of Theorem \ref{thm_model_shift}.
\begin{lemma}\label{lem_minv}[\todo{ref?}]
	Let $X\in\real^{n\times p}$ that contains i.i.d. row vectors with mean $0$ and covariance $\Sigma$.
	In the setting when $n = c p$ we have that as $p$ goes to infinity,
	\[ \bigtr{(X^{\top}X)^{-1}\Sigma} = \frac{1}{c - 1}. \]
\end{lemma}

\begin{lemma}\label{lem_cov_shift}
	\todo{restate the setting} In the setting of Theorem \ref{thm_model_shift}, we have with high probability $1-{\rm o}(1)$,
\begin{align}\label{lem_cov_shift_eq}
\tr ((X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2) = \frac{1}{n_1+n_2}\cdot \bigtr{ (a_1 M^\top M + a_2)^{-1}} +\bigo{ n^{-1/2+\epsilon}},
\end{align}
	for any constant $\epsilon>0$.
\end{lemma}
We will give the proof of Lemma \ref{lem_cov_shift} in Section \ref{sec_maintools}.

@@

Using Lemma \ref{lem_minv} and \ref{lem_cov_shift}, we can track the change of variance from $\hat{\beta}_t^{\MTL}$ to $\hat{\beta}_t^{\STL}$ as follows
The proof will consist of two main steps.
\begin{itemize}
	\item First, we show that $\hat{v}$ is close to $1$.
	\item Second, we plug $\hat{v}$ back into $\te(\hat{\beta}_t^{\MTL})$ to show the result.
\end{itemize}
